---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

**Note1: This is originally created by jennybc, you can find the original one [here](https://github.com/jennybc/foofactors)**
**NOTE2: This is a toy package created for expository purposes. It is not meant to actually be useful. If you want a package for factor handling, please see [forcats](https://cran.r-project.org/package=forcats).**

### foofactors

Factors are a very useful type of variable in R, but they can also drive you nuts. This package provides some helper functions for the care and feeding of factors.

### Installation

```{r installation, eval = FALSE}
devtools::install_github("STAT545-UBC-students/hw07-RyanGao67/foofactor")
```

### Package structure

- R scripts for functions are under ["R"](./R)
- Test scripts are under ["tests"](./tests)
- HTML generated by vignettes is ["hello-foofactors.html"](./doc/hello-foofactors.html)
  
### Quick demo

* Binding two factors via `fbind()`:

```{r}
library(foofactors)
a <- factor(c("character", "hits", "your", "eyeballs"))
b <- factor(c("but", "integer", "where it", "counts"))
```

* Simply catenating two factors leads to a result that most don't expect.

```{r}
c(a, b)
```

* The `fbind()` function glues two factors together and returns factor.

```{r}
fbind(a, b)
```

* Often we want a table of frequencies for the levels of a factor. The base `table()` function returns an object of class `table`, which can be inconvenient for downstream work. Processing with `as.data.frame()` can be helpful but it's a bit clunky.

```{r}
set.seed(1234)
x <- factor(sample(letters[1:5], size = 100, replace = TRUE))
table(x)
as.data.frame(table(x))
```

* The `freq_out()` function returns a frequency table as a well-named `tbl_df`:

```{r}
freq_out(x)
```


* `detect_factors()`

* The `fdetect()` function checks if a factor is actually a character, it basicly checks that if a factor has repeated value in which case the factor can be considered as a character

```{r}
# This is a factor
factor <- factor(c("m","i","s","s","i","s"))
# This is a character
character <- factor(c("s", "t", "a"))
                      
# return TRUE if it is a factor
detect_factors(factor)
# return FALSE if it is a character
detect_factors(character)
```

* `newreorder()`

* The `newreorder()` function can reorder the levels of a factor in descending order.

```{r}
# create a factor as an example
f <- factor(c("a", "b", "c","d"))
# the originally order: "a" "b" "c" "d"
levels(f)
# the reordered order "d" "c" "b" "a"
levels(newreorder(f))
```

* `newfactor()`
Set levels of a factor to the order in which they appear in the data.
* `newrev()`
Set levels of a factor to the reversed order in which they appear in the data.

```{r}
# create a factor as an example
f <- factor(c("b", "a", "c"))
# the originally order of level should be "b" "a" "c"
levels(f)
# the reset levels should be "b" "a" "c"
levels(newfactor(f))
# the reversed levels should be "c" "a" "b"
levels(newrev(f))
```

* `dfwrite()`/`dfread()`

Functions `dfwrite()` and `dfread()`: functions to write and read data frames to plain text delimited files while retaining factor levels

```{r}
# The basic idea here is that I first introduce a data frame
# Then I reorder the data
# Then I write the data frame to a file
# Then I read the data from the file
# Then I compare the data that I created and the data that I fetched from the file


dataframe <- data.frame(
    first = 1,
    second = 1:100,
    factor = sample(LETTERS[1:6], 10, replace = TRUE)
)

  dataframe$factor <- newreorder(dataframe$factor)

  dfwrite(dataframe, "./test_reorder.csv", "./levels_reorder.txt")
  fetched <- dfread("./test_reorder.csv", "./levels_reorder.txt")

levels(fetched$factor)
levels(dataframe$factor)

```


## References
- [Be the boss of your factors](https://www.stat.ubc.ca/~jenny/STAT545A/block08_bossYourFactors.html)
